\chapter{Revision History}

\section{Version 1.0 to Version 1.1}

Below are the changes incorporated into Version 1.1 of the \SBGNPDLone specification.

\begin{center}
%\tablecaption{Namespace scope definitions.}
\label{tab:revision history 1.1}
\tablefirsthead{\hline
Description & Tracker ID\\\hline}
\tablehead{\hline
\multicolumn{2}{|l|}{\small\slshape continued from previous page}\\
\hline\hline
Description & Tracker ID\\\hline\hline}
\tabletail{\hline
\multicolumn{2}{|r|}{\small\slshape continued on next page}\\
\hline}
\tablelasttail{\hline}
\begin{supertabular}{|p{12cm}|c|}\hline
Regarding modulation of reversible processes, changed ``should'' to ``must'' be represented by two \glyph{process} & \\\hline
Removed ``The connectors and the box move as a rigid entity'' in the definition of \glyph{process} & \\\hline
Changed the definition of process node to ``represent processes that transform one or several EPNs into \emph{one or several EPNs, identical or different}''& \\\hline
Changed SBO term of \glyph{compartment} From SBO:0000289 (functional compartment) to SBO:0000290 (physical compartment) & \\\hline
Reorganised classification of glyphs & \\\hline
Reorganised glyph section to reflect the above changes & \\\hline
Revised reference card to reflect changes in glyph organisation &
\\\hline
Revised logic operators throughout spec to make sure input and output
arcs meet before attaching to the glyph - as with processes. &
\\\hline
Added enumerated rules to grammar section. This is probably not
complete, but should help the implementation of semantic validation by
software tools. The hope is this will be refined as tools start
validating maps. & \\\hline
Updated UML maps and data dictionary  to be consistent with rest
of changes to spec. & \\\hline
Definition of cardinality is ambiguous & 2840996 \\\hline
\glyph{Sink and source} are lumped together & 2726435 \\\hline
SBO terms are incorrect or missing. & 2841261 \\\hline
\glyph{Compartment} description is confusing and contradictory. & 2841122
\\\hline
\glyph{Clone marker} fill percentages unhelpful. & 2841114 \\\hline
Use of CV for physical characteristic not clear. & 2841085 \\\hline
Definition of Cardinality is ambiguous. & 2840996 \\\hline
input to AND on IFN example. & 2804326 \\\hline
more SBO terms for \glyph{multimers} & 2803593 \\\hline
Legend of figure 2.20 is incorrect & 2803537 \\\hline
legend of figure 3.2 & 2802990 \\\hline
Compartment colouring & 2745703 \\\hline
Errors in diag a4. & 2664912 \\\hline
Change name of trigger glyph. & 2664908 \\\hline
Transition should be renamed process. & 2664862 \\\hline
Converting arcs tautological. & 2664843 \\\hline
Example invalid. & 2545870 \\\hline
consumption and production. & 2388317 \\\hline
Should require circles to be distinguishable from ellipses & 2219388
\\\hline
Figure 2.53 & 2162619 \\\hline
Reference card: production & 2104471 \\\hline
Figure 2.42 is wrong & 2104465 \\\hline
Mistake in the multi-cellular example  & 2395488 \\\hline
Should not prevent processes having identical in and out & 2664933
\\\hline
No description of linking to subunit rules. & 2545810 \\\hline
Extensively revised the grammar section. The UML diagrams have been simplified to show glyph taxonomy, and the data dictionary has been pruned to just show glyph identity. The some syntax rules have been moved into semantics and the rules reformulated to make them easier to understand. & \\\hline
Eliminated duplicate rules in layout section and revised text
slightly. & \\\hline
Phenotype cloning? & 2989007 \\\hline
Perturbing agent description & 2940021 \\\hline
\end{supertabular}
\end{center}

\section{Version 1.1 to Version 1.2}

Below are the changes incorporated into Version 1.2 of the \SBGNPDLone specification.

\begin{center}
%\tablecaption{Namespace scope definitions.}
\label{tab:revision history 1.2}
\tablefirsthead{\hline
Description & Tracker ID\\\hline}
\tablehead{\hline
\multicolumn{2}{|l|}{\small\slshape continued from previous page}\\
\hline\hline
Description & Tracker ID\\\hline\hline}
\tabletail{\hline
\multicolumn{2}{|r|}{\small\slshape continued on next page}\\
\hline}
\tablelasttail{\hline}
\begin{supertabular}{|p{12cm}|c|}\hline
Perturbing agent description & 2940021 \\\hline
Members of complex touching & 2849273 \\\hline
PD Reference card error for submap glyph & 3029242 \\\hline
\end{supertabular}
\end{center}

\section{Version 1.2 to Version 1.3}

Below are the changes incorporated into Version 1.3 of the \SBGNPDLone specification.

\begin{center}
%\tablecaption{Namespace scope definitions.}
\label{tab:revision history 1.3}
\tablefirsthead{\hline
Description & Tracker ID\\\hline}
\tablehead{\hline
\multicolumn{2}{|l|}{\small\slshape continued from previous page}\\
\hline\hline
Description & Tracker ID\\\hline\hline}
\tabletail{\hline
\multicolumn{2}{|r|}{\small\slshape continued on next page}\\
\hline}
\tablelasttail{\hline}
\begin{supertabular}{|p{12cm}|c|}\hline
Incorrect editor on title page & \\\hline
Typos in acknowledgements & \\\hline
Fixed typo in item on catalysis in section on modulation semantics. & \\\hline
State variables figure 2.6 V1.2 & 3090543 \\\hline\end{supertabular}
\end{center}

\section{Version 1.3 to Version 2.0}

The changes incorporated into Version 2.0 of the \SBGNPDLone are
summarised in chapter~\ref{chap:whatsnew}.
