\chapter{Acknowledgements}\label{sec:acknowledgements}

Here we acknowledge those people and organisations that assisted in the development of this and previous releases of the SBGN \PDl specification. First, we specifically acknowledge those who contributed directly to each revision of the  specification document, followed by a comprehensive acknowledgement of contributors that attended workshops and forum meetings or in some other way provided input to the standard. Finally, we acknowledge the bodies that provided financial support for the development of the standard.

% Here is a more comprehensive list of person who have been actively involved in SBGN development, either by their help designing the languages, their comments on the specification, help with development infrastructure or any other useful input.\\[\baselineskip]

%  We aim this list to be rather complete. We are very sorry if we forgot someone, and will be grateful if you notified us of any omission.

\section{Level 1 Release 1.0}

The specification of was written by Nicolas \lenov{}, 
Stuart Moodie, Anatoly Sorokin, Michael Hucka, Falk Schreiber, Emek Demir, 
Huaiyu Mi, Yukiko Matsuoka, Katja Wegner, and Hiroaki Kitano. In addition, 
the specification benefited much from the help of (in alphabetical order) Frank Bergmann, Sarala 
Dissanayake, Ralph Gauges, Peter Ghazal, Lu Li, and Steven Watterson.

% Mirit Aladjemm, Frank Bergmann, Emek Demir, Sarala Dissanayake, Ugur Dogrusoz, Tom Freeman, Akira Funahashi, Ralph Gauges, Peter Ghazal, Igor Goryanin, Michael Hucka, Akiya Jouraku, Sohyoung Kim, Hiroaki Kitano, Kurt Kohn, Fedor Kolpakov , Nicolas Le Nov\`{e}re, Lu Li, Yukiko Matsuoka, Huaiyu Mi, Stuart Moodie, Sven Sahle, Falk Schreiber, Anatoly Sorokin, Jessica Stephens, Linda Taddeo, Steven Watterson, Alice Villeger, Katja Wegner.\\[\baselineskip]

\section{Level 1 Release 1.1}

The specification of SBGN PD L1 V1.1 was written by Stuart Moodie and Nicolas \lenov{}, with contributions from (in alphabetical order) Frank Bergmann, Sarah Boyd, Emek Demir, Sarala Wimalaratne, Yukiko Matsuoka, Huaiyu Mi, Falk Schreiber, Anatoly Sorokin, and Alice Vill\'{e}ger.

\section{Level 1 Release 1.2}

The specification of SBGN PD L1 V1.2 was modified by Stuart Moodie, with contributions from (in alphabetical order) Sarah Boyd, Nicolas \lenov{}, and Huaiyu Mi.

\section{Level 1 Release 1.3}

The specification of SBGN PD L1 V1.3 was modified by Stuart Moodie, with contributions from (in alphabetical order) Tobias Czauderna, Nicolas \lenov{}, and Anatoly Sorokin.

\section{Level 1 Release 2.0}

The specification of SBGN PD L1 V2.0 was modified by Adrien Rougny, with contributions from (in alphabetical order) Irina Balaur, Michael Blinov, Hanna Borlinghaus, Ugur Dogrusoz, Andreas Dräger, Augustin Luna, Alexander Mazein, Stuart Moodie, and Vasundra Touré.


\section{Comprehensive list of acknowledgements}

Here is a more comprehensive list of people who have been actively involved in SBGN development, either by their help designing the languages, their comments on the specification, help with development infrastructure or any other useful input.  We intend this list to be complete. We are very sorry if we forgot someone, and would be grateful if you could notify us of any omission.

Mirit Aladjemm, Irina Balaur, Frank Bergmann, Michael Blinov, Hanna Borlinghaus, Sarah Boyd, Laurence Calzone, Melanie Courtot, Emek Demir, Ugur Dogrusoz, Andreas Dräger, Tom Freeman, Akira Funahashi, Ralph Gauges, Peter Ghazal, Samik Ghosh, Igor Goryanin, Michael Hucka, Akiya Jouraku, Hideya Kawaji, Douglas Kell, Sohyoung Kim, Hiroaki Kitano, Kurt Kohn, Fedor Kolpakov, Nicolas \lenov{}, Lu Li, Augustin Luna, Yukiko Matsuoka, Alexander Mazein, Huaiyu Mi, Stuart Moodie, Adrien Rougny, Sven Sahle, Chris Sander, Herbert Sauro, Esther Schmidt, Falk Schreiber, Jacky Snoep, Anatoly Sorokin, Jessica Stephens, Linda Taddeo, Vasundra Touré, Steven Watterson, Alice Vill\'{e}ger, Katja Wegner, Sarala Wimalaratne, Guanming Wu.


The authors are also grateful to all the attendees of the SBGN meetings, as 
well as to the subscribers of the \mailto{sbgn-discuss@googlegroups.com} mailing list.


\section{Financial Support}

The development of SBGN was mainly supported by a grant from the Japanese \emph{New Energy and Industrial Technology Development Organization} (NEDO, \url{http://www.nedo.go.jp/}).  The \emph{Okinawa Institute of Science and Technology} (OIST, \url{http://www.oist.jp/}), the \emph{AIST Computational Biology Research Center} (AIST CBRC, \url{http://www.cbrc.jp/index.eng.html}) the British \emph{Biotechnology and Biological Sciences Research Council} (BBSRC, \url{http://www.bbsrc.ac.uk/}) through a Japan Partnering Award, the European Media Laboratory (EML Research gGmbH, \url{http://www.eml-r.org/}), and the Beckman Institute at the California Institute of Technology (\url{http://bnmc.caltech.edu}) provided additional support for SBGN workshops. Some help was supplied by the \emph{Japan Science and Technology Agency} (JST, \url{http://www.jst.go.jp/}) and the \emph{Genome Network Project} of the Japanese Ministry of Education, Sports, Culture, Science, and Technology (MEXT, \url{http://www.mext.go.jp/}) for the development of the gene regulation network aspect of SBGN, and from the \emph{Engineering and Physical Sciences Research Council} (EPSRC, \url{http://www.epsrc.ac.uk}) during the redaction of the specification. The \emph{German Ministry of Education and Research} (\url{https://www.bmbf.de/}) provided support for an SBGN workshop and the redaction of the specification. Funding from \emph{National Institute of General Medical Sciences (NIGMS)} grant (P41 GM103504) and \emph{National Human Genome Research Institute (NHGRI)} grant (U41 HG006623) was additionally used for the redaction of the specification.
