\chapter{Revision History}

\section{Release 1.0 to Release 1.1}

Below are the changes incorporated into Release 1.1 of the \SBGNPDLone specification.

\begin{center}
%\tablecaption{Namespace scope definitions.}
\label{tab:revision history}
\tablefirsthead{\hline
Description & Tracker ID\\\hline}
\tablehead{\hline
\multicolumn{2}{|l|}{\small\sl continued from previous page}\\
\hline\hline
Description & Tracker ID\\\hline\hline}
\tabletail{\hline
\multicolumn{2}{|r|}{\small\sl continued on next page}\\
\hline}
\tablelasttail{\hline}
\begin{supertabular}{|p{12cm}|c|}\hline
Changed SBO term of \glyph{compartment} From SBO:0000289 (functional compartment) to SBO:0000290 (physical compartment) & \\\hline
Reoganised classification of glyphs & \\\hline
Reoganised glyph section to reflect the above changes & \\\hline
Revised reference card to reflect changes in glyph organisation &
\\\hline
Revised logic operators throughout spec to make sure input and output
arcs meet before attaching to the glyph - as with processes. &
\\\hline
Added enumerated rules to grammar section. This is probably not
complete, but should help the implementation of semantc validation by
software tools. The hope is this will be refined as tools start
validating maps. & \\\hline
Updated UML maps and data dictionary  to be consistent with rest
of changes to spec. & \\\hline
Definition of cardinality is ambiguous & 2840996 \\\hline
\glyph{Sink and source} are lumped together & 2726435 \\\hline
SBO terms are incorrect or missing. & 2841261 \\\hline
\glyph{Compartment} description is confusing and contradictory. & 2841122
\\\hline
\glyph{Clone marker} fill percentages unhelpful. & 2841114 \\\hline
Use of CV for physical charactetistic not clear. & 2841085 \\\hline
Definition of Cardinality is ambiguous. & 2840996 \\\hline
input to AND on IFN example. & 2804326 \\\hline
more SBO terms for \glyph{multimers} & 2803593 \\\hline
Legend of figure 2.20 is incorrect & 2803537 \\\hline
legend of figure 3.2 & 2802990 \\\hline
Compartment colouring & 2745703 \\\hline
Errors in diag a4. & 2664912 \\\hline
Change name of trigger glyph. & 2664908 \\\hline
Transition should be renamed process. & 2664862 \\\hline
Converting arcs tautological. & 2664843 \\\hline
Example invalid. & 2545870 \\\hline
consumption and production. & 2388317 \\\hline
Should require circles to be distinguishable from ellipses & 2219388
\\\hline
Figure 2.53 & 2162619 \\\hline
Reference card: production & 2104471 \\\hline
Figure 2.42 is wrong & 2104465 \\\hline
Mistake in the multi-cellular example  & 2395488 \\\hline
Should not prevent processes having identical in and out & 2664933
\\\hline
No description of linking to subunit rules. & 2545810 \\\hline
\end{supertabular}
\end{center}
