\chapter{Process Description Glyphs}
\label{chp:glyphs}

\section{Overview}

\input{pd_overview.tex}

\section{Controlled vocabularies used in \SBGNPDLone}\label{sec:CVs}

\input{cvs.tex}

\section{Auxiliary units}

Auxiliary units are glyphs that decorate other glyphs, providing additional information that may be useful to the reader.
In doing so, they change the meaning of the glyph or provide additional information about it.
These can provide specific annotation (\glyph{unit of information}), state information (\glyph{state variable}), indicate duplication of entity pool nodes (\glyph{clone marker}), describe specific glyphs (\glyph{subunit} for \glyph{complex}), or provide handles to elements lying outside of the maps (\glyph{submap terminal} for \glyph{submap}).

\input{unit_information.tex}
\input{state_variable.tex}
\input{clone_marker.tex}
\input{subunit.tex}
\input{submap_terminal.tex}

\section{Entity pool nodes}\label{sec:EPNs}

An entity pool is a population of entities that cannot be distinguished from each other when it comes to the \SBGNPDLone map.
For instance, all the molecular  entities that fulfill the same role in a given process form an entity pool.
As a result, an entity pool can represent different granularity levels, such as all the proteins, all the instances of a given protein, only certain forms of a given protein.
To belong to a different compartment is sufficient to belong to different entity pools.
Calcium ions in the endoplasmic reticulum and calcium ions in the cytosol belong to different entity pools when it comes to representing calcium release from the endoplasmic reticulum.

The \PD contains six glyphs representing classes of material entities: \glyph{unspecified entity} (\sect{unspecifiedEntity}), \glyph{simple chemical} (\sect{simpleChemical}), \glyph{macromolecule} (\sect{macromolecule}), \glyph{nucleic acid feature} (\sect{genetic}), \glyph{multimer} (\sect{multimer}) and \glyph{complex} (\sect{complex}).
(Specific types of macromolecules, such as protein, RNA, DNA, polysaccharide, and specific simple chemicals are not defined by \PD but may be part of future levels of SBGN.)
In addition to the material entities, \PD represents two conceptual entity pools: \glyph{empty set} (\sect{emptySet}), and \glyph{perturbing agent} (\sect{perturbing agent}).
Material and conceptual entities can optionally carry auxiliary units such as \glyph{units of information} (\sect{unitInfo}), \glyph{state variables}  (\sect{stateVariable}) and \glyph{clone markers} (\sect{cloneMarker}).

\input{unspecified.tex}
\input{simple_chemical.tex}
\input{macromolecule.tex}
\input{nucleic_acid_feature.tex}
\input{multimer.tex}
\input{complex.tex}
\input{empty_set.tex}
\input{perturbing_agent.tex}
\input{complex_epn_eg.tex}

\section{Defined sets of entity pool nodes}

\input{compartment.tex}

\section{Process nodes}\label{sec:PNs}

Process nodes represent processes that transform one or several entity pools into one or several entity pools, identical or different.  \SBGNPDLone defines a generic \glyph{process} (\sect{process}), as well as five more specific ones: the \glyph{omitted process} (\sect{omitted}), the \glyph{uncertain process} (\sect{uncertain}), the \glyph{association} (\sect{association}), the \glyph{dissociation} (\sect{dissociation}), and the \glyph{phenotype} (\sect{phenotype}).  In future levels of the SBGN \PDl, more processes may be defined.  (One can even envision the development of a controlled vocabulary of processes, as is done now for \glyph{EPNs}; see \sect{CVs}.)

\input{process.tex}
\input{omitted.tex}
\input{uncertain.tex}
\input{association.tex}
\input{dissociation.tex}
\input{phenotype.tex}

\section{Flux arcs}
\label{sec:fluxes}

Processes transform entity pools into other entity pools (\sect{PNs}).
Flux arcs allow representing which entity pools are consumed and produced by a process.
\glyph{Consumption} arcs link processes to their reactants, and \glyph{production} arcs link processes to their products.
\SBGNPDLone does not provide any specific arc to represent the fluxes of a reversible process, that can be conveniently represented using only \glyph{production} arcs.

\input{consumption.tex}
\input{production.tex}

\section{Modulation arcs}
\label{sec:modulations}

Modulation arcs represent influences of entity pools on processes.
A stimulation affects positively the flux of a process, while an inhibition affects it negatively.
\SBGNPDLone provides six modulation arcs: \glyph{modulation}, \glyph{stimulation}, \glyph{catalysis}, \glyph{necessary stimulation} and \glyph{inhibition}.

\input{modulation.tex}
\input{stimulation.tex}
\input{catalysis.tex}
\input{inhibition.tex}
\input{necessary_stim.tex}

\section{Logical operators}

\label{sec:logic}

The \glyph{logical operator} performs an operation on one or more inputs to give a unique output.
The \glyph{and}, \glyph{or}, and \glyph{not} operators perform a Boolean operation to give a binary output, while the \glyph{equivalence} operator performs a union of pools to give a new pool.

\input{and.tex}
\input{or.tex}
\input{not.tex}
\input{equivalence.tex}

\section{Logic arc}
\label{sec:logic_arc}

\input{logic_arc.tex}

\section{Annotating nodes and arcs}

\input{annotation.tex}

\section{Referring to other nodes}
\label{sec:ref_nodes}

Reference nodes handle links or relationships between elements of a map and sub-map.
At present there is only one reference glyph, \glyph{tag}, which can be used in a map referred to by a \glyph{submap} (\sect{submap}).


\input{tag.tex}
\input{equivalence_arc.tex}

\section{Encapsulation}
\label{sec:encapsulation}

\input{submap.tex}
