\chapter{Revision History}

\section{Version 1.0 to Version 1.1}

Below are the changes incorporated into Version 1.1 of the \SBGNPDLone specification.

\begin{center}
\label{tab:revision history 1.1}
\tablefirsthead{\hline
Description & Tracker ID\\\hline}
\tablehead{\hline
\multicolumn{2}{|l|}{\small\sl continued from previous page}\\
\hline\hline
Description & Tracker ID\\\hline\hline}
\tabletail{\hline
\multicolumn{2}{|r|}{\small\sl continued on next page}\\
\hline}
\tablelasttail{\hline}
\begin{supertabular}{|p{12cm}|c|}\hline
Regarding modulation of reversible processes, changed ``should'' to ``must'' be represented by two \glyph{process} & \\\hline
Removed ``The connectors and the box move as a rigid entity'' in the definition of \glyph{process} & \\\hline
Changed the definition of process node to ``represent processes that transform one or several EPNs into \emph{one or several EPNs, identical or different}''& \\\hline
Changed SBO term of \glyph{compartment} From SBO:0000289 (functional compartment) to SBO:0000290 (physical compartment) & \\\hline
Reoganised classification of glyphs & \\\hline
Reoganised glyph section to reflect the above changes & \\\hline
Revised reference card to reflect changes in glyph organisation &
\\\hline
Revised logic operators throughout spec to make sure input and output
arcs meet before attaching to the glyph - as with processes. &
\\\hline
Added enumerated rules to grammar section. This is probably not
complete, but should help the implementation of semantc validation by
software tools. The hope is this will be refined as tools start
validating maps. & \\\hline
Updated UML maps and data dictionary  to be consistent with rest
of changes to spec. & \\\hline
Definition of cardinality is ambiguous & 2840996 \\\hline
\glyph{Sink and source} are lumped together & 2726435 \\\hline
SBO terms are incorrect or missing. & 2841261 \\\hline
\glyph{Compartment} description is confusing and contradictory. & 2841122
\\\hline
\glyph{Clone marker} fill percentages unhelpful. & 2841114 \\\hline
Use of CV for physical charactetistic not clear. & 2841085 \\\hline
Definition of Cardinality is ambiguous. & 2840996 \\\hline
input to AND on IFN example. & 2804326 \\\hline
more SBO terms for \glyph{multimers} & 2803593 \\\hline
Legend of figure 2.20 is incorrect & 2803537 \\\hline
legend of figure 3.2 & 2802990 \\\hline
Compartment colouring & 2745703 \\\hline
Errors in diag a4. & 2664912 \\\hline
Change name of trigger glyph. & 2664908 \\\hline
Transition should be renamed process. & 2664862 \\\hline
Converting arcs tautological. & 2664843 \\\hline
Example invalid. & 2545870 \\\hline
consumption and production. & 2388317 \\\hline
Should require circles to be distinguishable from ellipses & 2219388
\\\hline
Figure 2.53 & 2162619 \\\hline
Reference card: production & 2104471 \\\hline
Figure 2.42 is wrong & 2104465 \\\hline
Mistake in the multi-cellular example  & 2395488 \\\hline
Should not prevent processes having identical in and out & 2664933
\\\hline
No description of linking to subunit rules. & 2545810 \\\hline
Extensively revised the grammar section. The UML diagrams have been simplified to show glyph taxonomy, and the data dictionary has been pruned to just show glyph identity. The some syntax rules have been moved into semantics and the rules reformulated to make them easier to understand. & \\\hline
Eliminated duplicate rules in layout section and revised text
slightly. & \\\hline
Phenotype cloning? & 2989007 \\\hline
Perturbing agent description & 2940021 \\\hline
\end{supertabular}
\end{center}

\section{Version 1.1 to Version 1.2}

Below are the changes incorporated into Version 1.2 of the \SBGNPDLone specification.

\begin{center}
\label{tab:revision history 1.2}
\tablefirsthead{\hline
Description & Tracker ID\\\hline}
\tablehead{\hline
\multicolumn{2}{|l|}{\small\sl continued from previous page}\\
\hline\hline
Description & Tracker ID\\\hline\hline}
\tabletail{\hline
\multicolumn{2}{|r|}{\small\sl continued on next page}\\
\hline}
\tablelasttail{\hline}
\begin{supertabular}{|p{12cm}|c|}\hline
Perturbing agent description & 2940021 \\\hline
Members of complex touching & 2849273 \\\hline
PD Reference card error for submap glyph & 3029242 \\\hline
\end{supertabular}
\end{center}

\section{Version 1.2 to Version 1.3}

Below are the changes incorporated into Version 1.3 of the \SBGNPDLone specification.

\begin{center}
\label{tab:revision history 1.3}
\tablefirsthead{\hline
Description & Tracker ID\\\hline}
\tablehead{\hline
\multicolumn{2}{|l|}{\small\sl continued from previous page}\\
\hline\hline
Description & Tracker ID\\\hline\hline}
\tabletail{\hline
\multicolumn{2}{|r|}{\small\sl continued on next page}\\
\hline}
\tablelasttail{\hline}
\begin{supertabular}{|p{12cm}|c|}\hline
Incorrect editor on title page & \\\hline
Typos in acknowledgements & \\\hline
Fixed typo in item on catalysis in section \ref{sec:mod-semantics}. & \\\hline
State variables figure 2.6 V1.2 & 3090543 \\\hline\end{supertabular}
\end{center}

\section{Version 1.3 to Version 2.0}

Below are the changes incorporated into Version 2.0 of the \SBGNPDLone specification.

\begin{center}
\label{tab:revision history 2.0}
\tablefirsthead{\hline
Description & Tracker ID\\\hline}
\tablehead{\hline
\multicolumn{2}{|l|}{\small\sl continued from previous page}\\
\hline\hline
Description & Tracker ID\\\hline\hline}
\tabletail{\hline
\multicolumn{2}{|r|}{\small\sl continued on next page}\\
\hline}
\tablelasttail{\hline}
\begin{supertabular}{|p{12cm}|c|}\hline
    Added equivalence operator glyph (\sect{equivalence}) & \\\hline
    Added annotation glyph (\sect{annotation}) & \\\hline
    Added submap terminal glyph (\sect{submapTerminal}) & \\\hline
    Replaced source and sink glyphs by empty set glyph (\sect{emptySet}) & \\\hline
    Added subunit glyphs (\sect{subunit}) & \\\hline
    Changed state variable glyph from ellipse to stadium (\sect{stateVariable}) & \\\hline
    Changed simple chemical glyph from circle to stadium (\sect{simpleChemical}) & \\\hline
    Set process glyph identity to ``instance'' (\sect{concepts}) & \\\hline
    Broke arcs section into three new sections: flux arcs (\sect{fluxes}), modulation arcs (\sect{modulations}) and logic arc (\sect{logic_arc}) & \\\hline
    Moved sections ``Referring to other nodes and arcs'' (\sect{ref_nodes}) and ``Encapsulation'' (\sect{encapsulation}) to the end of chapter 2 & \\\hline
    Homogenized ``Container'', ``Label'' and ``Auxiliary units'' entries for all glyphs & \\\hline
    Redrew all figures with SBGN-ED using a unified style & \\\hline
    Added ``Incoming arcs'' and ``Outgoing arcs'' entries for all nodes & \\\hline
    Updated the reference card with the new glyphs and shapes (appendix~\ref{chap:refcard}) & \\\hline
    Added acknowledgements (\sect{acknowledgements}) & \\\hline
    Rewrote the standfirst and the ``Label'' entry of the state variable section (\sect{stateVariable}) & \\\hline
    Added text to the standfirst of the multimer section and table 2.8 showing the different glyphs (\sect{multimer}) & \\\hline
    Rewrote standfirst of the tag section (\sect{tag}) & \\\hline
    Rewrote standfirst of the submap section, some text explaining the examples and captions of the figures (\sect{submap}) & \\\hline
    Moved equivalence arc section (\sect{equivalenceArc}) to section ``Encapsulation'' (\sect{encapsulation}) and rewrote the standfirst & \\\hline
    Added a standfirst to the logical operators section (\sect{logic}) & \\\hline
    Rewrote the definitions of the and operator (\sect{and}), the or operator (\sect{or}) and the not operator (\sect{not}) glyphs & \\\hline
    Updated figure 3.1, table 3.1, table 3.2, and the tables of section 3.4.1 and 3.4.2 with the new glyphs & \\\hline
    Added rules regarding the equivalence operator glyph in the semantic rules section (\sect{eq-semantics}) & \\\hline
    Added usage examples for the equivalence operator glyph in a new appendix (appendix~\ref{ex_eq}) & \\\hline
\end{supertabular}
\end{center}
