% $HeadURL$

% =============================================================================
% introduction
% =============================================================================

\chapter{Introduction}

With the rise of systems and synthetic biology, the use of graphical representations
of pathways and networks to describe biological systems has become pervasive. 
It was therefore important to use a consistent notation that would allow people 
to interpret those maps easily and quickly, without the need of extensive legends.
Futhermore, distributed investigation of biological systems in different labs as well
as activities like synthetic biology, that reconstruct biological systems, need
to exchange their descriptions unambiguously, as engineers exchange circuit diagrams.
The goal of the Systems Biology Graphical Notation (SBGN) is to standardize the 
graphical/visual representation of biochemical and cellular processes. SBGN 
defines comprehensive sets of symbols with precise semantics, together with 
detailed syntactic rules defining their use. It also describes the manner 
in which such graphical information should be interpreted.
SBGN is made up of three different and complementary languages \cite{Novere:2009p1}.
This document defines the \emph{\PD{}} visual language of SBGN. \PDs are 
one of three views of a biological process offered by SBGN. It is the 
product of many hours of discussion and development by many individuals and 
groups.

\section{What are the languages?}
\label{sec:what-languages}

\textbf{PD} is a language that permits the description of all the processes
taking place in a biological system. The ensemble of all these processes constitute
a Description. \textbf{ER} is a language that permits the description of all 
the relations involving the entities of a biological system. The ensemble of all 
these relations constitute a Relationship. \textbf{AF} is a language that permits the 
description of the flow of activity in a biological system.

\section{Nomenclature}
\label{sec:nomenclature}

The three languages of SBGN should be referred to as:
\begin{compactitem}
\item the Process Description language.
\item the Entity Relationship language.
\item the Activity Flow language.
\end{compactitem}

Abbreviated as:
\begin{compactitem}
\item the PD language.
\item the ER language.
\item the AF language.
\end{compactitem}

A specific representation of a biological system in one of the SBGN languages should be referred to as:
\begin{compactitem}
\item a Process Description map.
\item an Entity Relationship map.
\item an Activity Flow map.
\end{compactitem}

Abbreviated as:
\begin{compactitem}
\item a PD map.
\item an ER map.
\item an AF map.
\end{compactitem}

The corpus of all SBGN representations should be referred to as:
\begin{compactitem}
\item Process Descriptions.
\item Entity Relationships.
\item Activity Flows.
\end{compactitem}

The capitalization is important. PD, ER and AF are names of languages. 
As such they must be capitalized in English. This is not the case of the accompanying noun (language or map).



\section{SBGN levels and versions}
\label{sec:sbgn-levels}

It was clear at the outset of SBGN development that it would be impossible 
to design a perfect and complete notation right from the beginning.  Apart 
from the prescience this would require (which, sadly, none of the authors 
possess), it also would likely need a vast language that most newcomers 
would shun as being too complex.  Thus, the SBGN community followed an idea 
used in the development of other standards, i.e. stratify language 
development into levels.

A \emph{level} of one of the SBGN languages represents a set of features 
deemed to fit together cohesively, constituting a usable set of 
functionality that the user community agrees is sufficient for a reasonable 
set of tasks and goals. Within \emph{levels}, \emph{versions} represent 
small evolution of a language, that may involve new glyphs, refined 
semantics, but no fundamental change of the way maps are to be generated 
and interpreted. In addition new versions should be backwards compatible, 
\ie \PD maps that conform to an earlier version of the \PDl within 
the same level should still be valid. This does not apply to a new level.
Capabilities and features that cannot be agreed upon and 
are judged insufficiently critical to require inclusion in a given level, 
are postponed to a higher level or version.  In this way, the development 
of SBGN languages is envisioned to proceed in stages, with each higher 
levels adding richness compared to the levels below it.

\section{Developments, discussions, and notifications of updates}
\label{sec:discussions}

The SBGN website (\url{http://sbgn.github.io/sbgn/}) is a portal for all things 
related to SBGN.  It provides a web forum interface to the SBGN discussion 
list (\mailto{sbgn-discuss@caltech.edu}) and information about how anyone 
may subscribe to it.  The easiest and best way to get involved in SBGN 
discussions is to join the mailing list and participate.

Face-to-face meetings of the SBGN community are announced on the website as 
well as the mailing list.  Although no set schedule currently exists for 
workshops and other meetings, we envision holding at least one public 
workshop per year.  As with other similar efforts, the workshops are likely 
to be held as satellite workshops of larger conferences, enabling attendees 
to use their international travel time and money more efficiently.

Notifications of updates to the SBGN specification are also broadcast on 
the mailing list and announced on the SBGN website.

\section{Note on typographical convention}
The concept represented by a glyph is written using a normal font, while a 
\glyph{glyph} means the SBGN visual representation of the concept. For 
instance ``a biological process is encoded by the SBGN PD \glyph{process}''.
A UML class is written in camel case and presented as CamelCaseClassName.
\textcolor{red}{Attributes are written in lower case and presented as camelCaseAttribute.}


\textcolor{red}{Note that in drafts of this document the margin-note is used to indicate that this
part of the document may require special attention from the reader as it may contain
material that is new or potentially contentious.}

% The following is for [X]Emacs users.  Please leave in place.
% Local Variables:
% TeX-master: "../sbgn_PD-level1"
% End:
